A re-examination of the role of friction in the original Social Force Model

### 2019-05-27 paper.tex
cambio de prueba
termine de agregar los ultimos comentarios (falta algunos que discutire con guille)

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics have been developed. One of the first models was developed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models represent the interaction between 
pedestrians and their environment using different kinds of forces inspired in physical systems. For 
example, Okazaki (1979) modeled the movement the pedestrians as the motion 
of a magnetized object within a magnetic field \cite{Okazaki}.}\\

\textcolor{red}{By the late '90s and the beginning of the century, Helbing 
\textit{et al.} proposed a model that included socio-psychological and physical 
forces to simulate crowd dynamics. According to Helbing \textit{et al.}, both 
the environment and the individuals' own desire affect the pedestrians motion in 
a similar way as forces do with respect to the momentum of particles 
\cite{Helbing1,Helbing4}. This ``Social Force Model'' (SFM) related the 
socio-psychological phenomenon of crowds behavior to the ``microscopic'' 
formalism of moving particles. The model succeeded at this instance to 
reproduce the slow down in the efficiency in the evacuation process as 
pedestrians try harder to escape from a dangerous situation (\textit{i.e.} 
``faster is slower'' effect) \cite{Helbing1,Dorso1,Dorso2}.} \\

@@

It is well known that the \textcolor{red}{original} version of the Social Force 
Model \textcolor{red}{\cite{parisi2}} is incapable of reproducing the 
fundamental diagram at high densities (say above 
\textcolor{red}{5~p~m$^{-2}$)}. Different 
approaches were proposed to fix this drawback: increasing the net-time 
headway~\cite{johansson}, canceling the desired velocity~\cite{parisi2} or even 
inducing the jamming state by an attraction target~\cite{kwak}. These approaches 
seems suitable when individuals are not so anxious to reach a certain 
destination. However, when individuals are escaping or rushing due to a 
stressful situation, pedestrians would neither consider reducing their desired 
force nor being captured by an attraction. \\

@@

\subsection*{\label{software}Simulation software}

\textcolor{red}{The simulations were implemented on LAMMPS molecular dynamics 
simulator with parallel computing capabilities \cite{plimpton}. The time 
integration algorithm followed the velocity Verlet scheme with a time step of 
$10^{-4}$~s. All the necessary parameters were set to the same values as in 
previous works (see Refs. \cite{Dorso5,sticco}), except for the friction 
coefficient $\kappa$.} \\

\textcolor{red}{We implemented special modules in C++ for upgrading the LAMMPS 
capabilities to attain the Social Force Model simulations. We also checked over 
the LAMMPS output with previous computations (see Refs. \cite{Dorso1, 
Dorso2,Dorso3, Dorso4,Dorso6}).}\\

\textcolor{red}{The measurements were taken once the system reached the 
stationary state ($t=30$~s), while the configurations of the systems were 
recorded every 0.05~s, that is, at intervals as short as 10\% of the 
pedestrian’s relaxation time (see Section. \ref{sfm}). The recorded magnitudes 
were the pedestrian’s positions and velocities for each process. We also 
computed the clustering structures using a LAMMPS built in function 
named compute cluster-atom. This function assigns each pedestrian a cluster ID.
A cluster is defined as a set of pedestrians, each of which is within the cutoff
distance from one or more other pedestrians in the cluster. 
If a pedestrian has no neighbors within the cutoff distance, then it is a 1-pedestrian cluster.}\\

@@

We show that the experimental data (say, the flux-density diagram) can be 
modeled under quasi-stationary conditions in the high density regime. Our 
starting point is the re-examination of the (basic) SFM.
\textcolor{red}{We proposed to reduce in units the equation of motion of 
the SFM (Eq.~\ref{eqn_appendix_1} and Eq.~\ref{eqn_appendix_2}). This leads us
to two control parameters} $\mathcal{A}$ and $\mathcal{K}$ (see Section~\ref{appendix_1}) may handle
the collective motion in dense crowds. We presume that physical contact is
a key feature in dense crowds, despite the fact that other issues may
also contribute to the flow reduction \cite{johansson1}. However, the latter could
be satisfactorily omitted in past research \cite{johansson}, and thus, we will not
attempt to introduce further extensions to the (basic) SFM for the
sake of simplicity.\\

@@

We found two unexpected results. Increasing the density produces bigger size clusters until the size distribution suddenly switches to a bimodal distribution (compare \ref{size_distribution_w22_density5} with \ref{size_distribution_w22_density5_5} and \ref{size_distribution_w22_density4.5_kx10} with \ref{size_distribution_w22_density5_kx10}). Once
this phenomenon occurs, any of two possibilities may
appear: the pedestrian belongs to a small cluster (out
of many caged in the crowd) or he (she) belongs to the
giant cluster (with a size comparable to the entire crowd).\\

@@

\bibitem{Boltes}
\textcolor{red}{Boltes M., Zhang J., Tordeux A., Schadschneider A., Seyfried A., 
2018. Empirical Results of Pedestrian and Evacuation Dynamics. In: Meyers R. 
(eds) Encyclopedia of Complexity and Systems Science. Springer, Berlin, 
Heidelberg}

@@

\bibitem{Baqui1}
\textcolor{red}{Baqui, Muhammad, and Rainald Löhner. "Real-time crowd safety
and comfort management from CCTV images." Real-Time Image and Video Processing 2017.
Vol. 10223. International Society for Optics and Photonics, 2017.}

\bibitem{Togashi1}
\textcolor{red}{Togashi, Fumiya, et al. "Using ensemble Kalman filter to
determine parameters for computational crowd dynamics simulations." 
Engineering Computations 35.7 (2018): 2612-2628.}

\bibitem{lohner2}
\textcolor{red}{L\"ohner, Rainald. "On the modeling of pedestrian motion." 
Applied Mathematical Modelling 34.2 (2010): 366-382.}

@@

\bibitem{kwak}
Kwak, Jaeyoung, Jo, Hang-Hyun, Luttinen, T., and Kosonen, L., 2017. "Jamming 
transitions induced by an attraction in pedestrian flow." Physical Review E 96 
(2), 022319.
{\path{https://doi.org/10.1103/PhysRevE.96.022319}}

@@

\bibitem{guo1}
Guo, Wei, Xiaolu Wang, and Xiaoping Zheng., 2015. "Lane formation in pedestrian 
counterflows driven by a potential field considering following and avoidance 
behaviours." Physica A: Statistical Mechanics and its Applications 432, 87-101.
{\path{https://doi.org/10.1016/j.physa.2015.03.020}}

\bibitem{qiao1}
Qiao, Jing and Sun, Lishan and Qiu, Shi and Rong, Jian and Liu, Xiaoming, 2017. 
"Reducing Bidirectional Pedestrian Conflict Based on Lane Formation Phenomenon 
in Subway Corridors." Promet-Traffic\&Transportation 29 (5), 489-502.
{\path{https://doi.org/10.7307/ptt.v29i5.2168}}

@@

\bibitem{lin}
Lin, Peng, Jian Ma, and Siuming Lo., 2016. "Discrete element crowd model for 
pedestrian evacuation through an exit." Chinese Physics B 25 (3), 034501.
{\path{https://doi.org/10.1088/1674-1056/25/3/034501}}

@@

\bibitem{li1}
Meifang Li, Yongxiang Zhao, Lerong He, Wenxiao Chen and Xianfeng Xu, 
2015. "The parameter calibration and optimization of social force model for the 
real-life 2013 Ya’an earthquake evacuation in China." Safety science 79, 
243-253.
{\path{https://doi.org/10.1016/j.ssci.2015.06.018}}

>>>>

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics have been developed. The first model was developed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models represent the interaction between 
pedestrians and their environment using different kinds of forces inspired in physical systems. For 
example, Okazaki (1979) modeled the movement the pedestrians as the motion 
of a magnetized object within a magnetic field \cite{Okazaki}.}\\

By the late '90s and the beginning of the century, Helbing 
\textit{et al.} proposed a model that included socio-psychological and physical 
forces to simulate crowd dynamics. According to Helbing \textit{et al.}, both 
the environment and the individuals' own desire affect the pedestrians motion in 
a similar way as forces do with respect to the momentum of particles 
\cite{Helbing1,Helbing4}. This ``Social Force Model'' (SFM) related the 
socio-psychological phenomenon of crowds behavior to the ``microscopic'' 
formalism of moving particles. The model succeeded at this instance to 
reproduce the slow down in the efficiency in the evacuation process as 
pedestrians try harder to escape from a dangerous situation (\textit{i.e.} 
``faster is slower'' effect) \cite{Helbing1,Dorso1,Dorso2}. \\

@@

It is well known that the \textcolor{red}{original} version of the Social Force 
Model  is incapable of reproducing the 
fundamental diagram at high densities (say above 
\textcolor{red}{5~p~m$^{-2}$)} \textcolor{red}{\cite{parisi2}} . Different 
approaches were proposed to fix this drawback: increasing the net-time 
headway~\cite{johansson}, canceling the desired velocity~\cite{parisi2} or even 
inducing the jamming state by an attraction target~\cite{kwak}. These approaches 
seems suitable when individuals are not so anxious to reach a certain 
destination. However, when individuals are escaping or rushing due to a 
stressful situation, pedestrians would neither consider reducing their desired 
force nor being captured by an attraction. \\

@@

\subsection{\label{software}\textcolor{red}{Simulation software}}

The simulations were implemented on LAMMPS molecular dynamics 
simulator with parallel computing capabilities \cite{plimpton}. The time 
integration algorithm followed the velocity Verlet scheme with a time step of 
$10^{-4}$~s. All the necessary parameters were set to the same values as in 
previous works (see Refs.~\cite{Dorso5,sticco}), except for the friction 
coefficient $\kappa$. \\

We implemented special modules in C++ for upgrading the LAMMPS 
capabilities to attain the Social Force Model simulations. We also checked over 
the LAMMPS output with previous computations (see Refs. \cite{Dorso1, 
Dorso2,Dorso3, Dorso4,Dorso6}).\\

The measurements were taken once the system reached the 
stationary state ($t=30$~s), while the configurations of the systems were 
recorded every 0.05~s, that is, at intervals as short as 10\% of the 
pedestrian’s relaxation time (see Section. \ref{sfm}). The recorded magnitudes 
were the pedestrian’s positions and velocities for each process. We also 
computed the clustering structures using a LAMMPS built-in function 
named compute cluster-atom. This function assigns each pedestrian a cluster ID.
A cluster is defined as a set of pedestrians, each of which is within the cutoff
distance from one or more other pedestrians in the cluster. 
If a pedestrian has no neighbors within the cutoff distance, then it is a 1-pedestrian cluster.\\

@@

We show that the experimental data (say, the flux-density diagram) can be 
modeled under quasi-stationary conditions in the high density regime. Our 
starting point is the re-examination of the (basic) SFM.
\textcolor{red}{We proposed to reduce in units the equation of motion of 
the SFM (Eq.~\ref{eqn_appendix_1} and Eq.~\ref{eqn_appendix_2}). This leads us
to two control parameters} $\mathcal{A}$ and $\mathcal{K}$ (see Section~\ref{appendix_1}) that may handle
the collective motion in dense crowds. We presume that physical contact is
a key feature in dense crowds, despite the fact that other issues may
also contribute to the flow reduction \cite{johansson1}. However, the latter could
be satisfactorily omitted in past research \cite{johansson}, and thus, we will not
attempt to introduce further extensions to the (basic) SFM for the
sake of simplicity.\\

@@

We found two unexpected results. Increasing the density produces bigger size clusters until the size distribution suddenly switches to a bimodal distribution (compare \ref{size_distribution_w22_density5} with \ref{size_distribution_w22_density5_5} and \ref{size_distribution_w22_density4.5_kx10} with \ref{size_distribution_w22_density5_kx10}). Once
this phenomenon occurs, any of two possibilities may
appear: the pedestrian belongs to a small cluster (out
of many caged in the crowd) or he (she) belongs to the
giant cluster (with a size comparable to the entire crowd).
\textcolor{red}{In other words, the bimodal distribution occurs because when 
the giant component is formed there are some small clusters ``caged"
inside the giant component. These small clusters do not touch any 
pedestrian of the giant component.} \\

@@

\bibitem{Boltes}
\textcolor{red}{Boltes, M., Zhang, J., Tordeux, A., Schadschneider A., Seyfried A., 
2018. Empirical Results of Pedestrian and Evacuation Dynamics. In: Meyers R. 
(eds) Encyclopedia of Complexity and Systems Science. Springer, Berlin, 
Heidelberg}

@@

\bibitem{Baqui1}
\textcolor{red}{Baqui, M., and L\"ohner, R. "Real-time crowd safety
and comfort management from CCTV images." Real-Time Image and Video Processing 2017.
Vol. 10223. International Society for Optics and Photonics, 2017.}

\bibitem{Togashi1}
\textcolor{red}{Togashi, F., Misaka, T., L\"ohner, R., \& Obayashi, S. (2018). 
Using ensemble Kalman filter to determine parameters for computational crowd dynamics simulations. 
Engineering Computations, 35(7), 2612-2628.}

\bibitem{lohner2}
\textcolor{red}{L\"ohner, R. "On the modeling of pedestrian motion." 
Applied Mathematical Modelling 34.2 (2010): 366-382.}

@@

\bibitem{kwak}
Kwak, J, Hang-Hyun, J., Luttinen, T., and Kosonen, L., 2017. "Jamming 
transitions induced by an attraction in pedestrian flow." Physical Review E 96 
(2), 022319.
{\path{https://doi.org/10.1103/PhysRevE.96.022319}}

@@

\bibitem{guo1}
Wei, G., Wang X., and Zheng, X., 2015. "Lane formation in pedestrian 
counterflows driven by a potential field considering following and avoidance 
behaviours." Physica A: Statistical Mechanics and its Applications 432, 87-101.
{\path{https://doi.org/10.1016/j.physa.2015.03.020}}

\bibitem{qiao1}
Jing, Q., Lishan, S., Shi, Q., Jian, R.  and Xiaoming L., 2017. 
"Reducing Bidirectional Pedestrian Conflict Based on Lane Formation Phenomenon 
in Subway Corridors." Promet-Traffic\&Transportation 29 (5), 489-502.
{\path{https://doi.org/10.7307/ptt.v29i5.2168}}

@@

\bibitem{lin}
Peng, L., Ma, J. and Lo, S. 2016. "Discrete element crowd model for 
pedestrian evacuation through an exit." Chinese Physics B 25 (3), 034501.
{\path{https://doi.org/10.1088/1674-1056/25/3/034501}}

@@

\bibitem{li1}
Li, M., Zhao, Y/., He, L., Chen, W. and Xu, X. 
2015. "The parameter calibration and optimization of social force model for the 
real-life 2013 Ya’an earthquake evacuation in China." Safety science 79, 
243-253.
{\path{https://doi.org/10.1016/j.ssci.2015.06.018}}
